### 2013-12-09 presentationPTD.tex
add presentation, remove old .Rproj

% GRADUATE INSTITUTE, GENEVA
% Bastiaan Quast
% bastiaan.quast@graduateinstitute.ch

\documentclass{beamer}

\mode<presentation>
{
  \usetheme{Berkeley}  
  \setbeamercovered{transparent}
}

\usecolortheme[colortext]{geneva}

\usenavigationsymbolstemplate{}

\usepackage[english]{babel}
\usepackage[utf8]{inputenc}
\usepackage{times}
\usepackage[T1]{fontenc}

\title[]{Preliminary Thesis Defence}
\subtitle{December 18, 2013}

% Supervisor:&\footnote{Professor of Economics, The Graduate Institute, Geneva; }\\
% Second Reader:&\footnote{Professor of Economics, The Graduate Institute, Geneva; }

\author[Bastiaan Quast]{
  Jean-Louis Arcand\inst{1,4}
    \and
  Ugo Panizza\inst{2,4}
    \and
  \\Bastiaan Quast\inst{3,4}
  }

\institute[The Graduate Institute, Geneva] % (optional, but mostly needed)
{
  \inst{1}
  Professor of Economics,\\
  Supervisor
    \and
  \inst{2}
  Professor of Economics,\\
  Second Reader
    \and
  \inst{3}
  PhD Student,\\
  bastiaan.quast@graduateinstitute.ch
    \and
  \inst{4}
  The Graduate Institute, Geneva
}

\date[18-12-2013]%{December 18, 2013}

\AtBeginSubsection[]
{
  \begin{frame}<beamer>{Outline}
    \tableofcontents[currentsection,currentsubsection]
  \end{frame}
}

\begin{document}

\begin{frame}{PhD Development Economics}
  \titlepage
\end{frame}

\section{Introduction}

\begin{frame}{Introduction}
  \tableofcontents
  % You might wish to add the option [pausesections]
\end{frame}

\section[Pensions, Child Growth]{Pensions and Child Growth: Additional Evidence from South-Africa}

\begin{frame}{Pensions and Child Growth}{Additional Evidence from South Africa}
using the new data
\end{frame}

\section[Surviving Andersonville]{Replication: Surviving Andersonville}

\begin{frame}{Replication: Surviving Andersonville}
these guys mislabeled everything
\end{frame}

\section[Uncertainty, Currency Attacks]{Uncertainty and Risk in Currency Attacks}

\begin{frame}{Uncertainty and Risk in Currency Attacks}{Knightian Uncertainty in Morris and Shin (1998)}
This paper makes no sense!!
\end{frame}

\section[Bitcoin, Stupid Phones]{Bitcoin and Remittances: Stupid Phone Potential}

\begin{frame}{Bitcoin and Remittances}{Stupid Phone Potential}
stupid phones rock!!
\end{frame}

\section{End Notes}

\begin{frame}{End Notes}
  \begin{itemize}
  \item
    The \alert{first main message} of your talk in one or two lines.
  \item
    The \alert{second main message} of your talk in one or two lines.
  \item
    Perhaps a \alert{third message}, but not more than that.
  \end{itemize}

  % The following outlook is optional.
  \vskip0pt plus.5fill
  \begin{itemize}
  \item
    Outlook
    \begin{itemize}
    \item
      Something you haven't solved.
    \item
      Something else you haven't solved.
    \end{itemize}
  \end{itemize}
\end{frame}

\end{document}